```{r setup, include=FALSE}
# set global chunk options
library(pacman); p_load(knitr, RWordPress, xtable, reports, knitcitations, animation)
opts_chunk$set(cache = TRUE, warning = FALSE)
bib <- read.bibtex(dir()[tools::file_ext(dir()) == "bib"][1])


pressNum_n_dir <- "61803503/wp/"
path2figureStorage <- "C:/Users/trinker/Dropbox/Public/wp/"
opts_knit$set(base.url = paste0("http://dl.dropboxusercontent.com/u/", pressNum_n_dir),
    base.dir = path2figureStorage)
opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE, tidy=FALSE)
library(knitr)
```

In this post you will learn how to:

1. Create your own quasi-shape file   
2. Plot your homemade quasi-shape file in `ggplot2`
3. Add an external svg/ps graphic to a plot   
4. Change a `grid` grob's color and alpha    

---

## Background <font size="2.5">(See [just code](https://raw.githubusercontent.com/trinker/space_manikin/master/space_manikin.R) if you don't care much about the process) </font>    

I started my journey wanting to replicate a graphic called a *space manikin* by `r citet(bib[["McNeil2005"]])` and fill areas in that graphic like a choropleth.  I won't share the image from McNeil's book as it's his intellectual property but know that the graphic is from a gesturing book that divides the body up into zones (p. 275).  To get a sense of what the manikin looks like here is the `ggplot2` version of it:

<div class="figure">
`r IM("https://raw.githubusercontent.com/trinker/space_manikin/master/images/manikin1.png", center = FALSE, width=300, height=300)`
`r caption_figure("fig1", "ggplot2 Version of McNeil's (2005) Space Manikin")`
</div>

While this is a map of areas of a body you can see where this could be extended to any number of spatial tasks such as mapping the layout of a room.

---

## 1. Creating a Quasi-Shape File

So I figured "zones" that's about like states on a map.  I have toyed with [choropleth maps](http://trinkerrstuff.wordpress.com/2013/07/05/ggplot2-chloropleth-of-supreme-court-decisions-an-tutorial/) of the US in the past and figured I'd generalize this learning.  The difference is I'd have to make the shape file myself as the `r HR("maps", "http://cran.r-project.org/web/packages/maps/index.html")` package doesn't seem to have McNeil's space manikin.

Let's look at what `ggplot2` needs from the `maps` package:

```{r}
library(maps); library(ggplot2)
head(map_data("state"))
```

Hmm coordinates, names of regions, and order to connect the coordinates.  I figured I can handle that.  I don't 100% know what a shape file is, mostly that it's a file that makes shapes.  What we're making may or may not technically be a shape file but know we're going to map shapes in `ggplot2` (I use the quasi to avoid the wrath of those who do know precisely what a shape file is).  

I needed to make the zones around an image of a person so I first grabbed a free png silhouette from: http://www.flaticon.com/free-icon/standing-frontal-man-silhouette_10633.  I then knew I'd need to add some lines and figure out the coordinates of the outlines of each cell.  So I read the raster image into R, plotted in `ggplot2` and added lots of grid lines for good measure.  Here's what I wound up with:

```{r, results='hide', fig.height=4.5, fig.width=4.5}
library(png); library(grid); library(qdap)
url_dl(url="http://i.imgur.com/eZ76jcu.png")
file.rename("eZ76jcu.png", "body.png")
img <- rasterGrob(readPNG("body.png"), 0, 0, 1, 1, just=c("left","bottom"))
ggplot(data.frame(x=c(0, 1), y=c(0, 1)), aes(x=x, y=y)) + 
	geom_point() +
	annotation_custom(img, 0, 1, 0, 1) + 
	scale_x_continuous(breaks=seq(0, 1, by=.05))+ 
	scale_y_continuous(breaks=seq(0, 1, by=.05)) + theme_bw() +
	theme(axis.text.x=element_text(angle = 90, hjust = 0, vjust=0))
```
`r caption_figure("fig2", "Silhouette from ggplot2 With Grid Lines")`

---

### 1b. Dirty Deeds Done Cheap

I needed to get reference lines on the plot so I could begin recording coordinates.  Likely there's a better process but this is how I approached it and it worked.  I exported the ggplot in `r HR("#fig2", "Figure 2")` into ([GASP](http://www.dramabutton.com/)) Microsoft Word (I may have just lost a few die hard command line folks).  I added lines and figured out the coordinates of the lines.  It looked something like this:

<div class="figure">
`r IM("https://raw.githubusercontent.com/trinker/space_manikin/master/images/manikin.png", center = FALSE, width=500, height=300)`
`r caption_figure("fig3", "Silhouette from ggplot2 with MS Word Augmented Border Lines")`
</div>

After that I began the tedious task of figuring out the corners of each of the shapes ("zones") in the space manikin.  Using `r HR("#fig3", "Figure 3")` and a list structure in R I mapped each of the corners, the approximate shape centers, and the order to plot the coordinates in for each shape.  This is the code for corners:


```{r}
library(qdap)
dat <- list(
    `01`=data.frame(x=c(.4, .4, .6, .6), y=c(.67, .525, .525, .67)),
    `02`=data.frame(x=c(.35, .4, .6, .65), y=c(.75, .67, .67, .75)),
    `03`=data.frame(x=c(.6, .65, .65, .6), y=c(.525, .475, .75, .67)),
    `04`=data.frame(x=c(.4, .35, .65, .6), y=c(.525, .475, .475, .525)),
    `05`=data.frame(x=c(.35, .35, .4, .4), y=c(.75, .475, .525, .67)),
    `06`=data.frame(x=c(.4, .4, .6, .6), y=c(.87, .75, .75, .87)),
    `07`=data.frame(x=c(.6, .6, .65, .65, .73, .73), y=c(.87, .75, .75, .67, .67, .87)),
    `08`=data.frame(x=c(.65, .65, .73, .73), y=c(.67, .525, .525, .67)),
    `09`=data.frame(x=c(.6, .6, .73, .73, .65, .65), y=c(.475, .28, .28, .525, .525, .475)),
    `10`=data.frame(x=c(.4, .4, .6, .6), y=c(.475, .28, .28, .475)),
    `11`=data.frame(x=c(.27, .27, .4, .4, .35, .35), y=c(.525, .28, .28, .475, .475, .525)),
    `12`=data.frame(x=c(.27, .27, .35, .35), y=c(.67, .525, .525, .67)),
    `13`=data.frame(x=c(.27, .27, .35, .35, .4, .4), y=c(.87, .67, .67, .75, .75, .87)),
    `14`=data.frame(x=c(.35, .35, .65, .65), y=c(1, .87, .87, 1)),
    `15`=data.frame(x=c(.65, .65, .73, .73, 1, 1), y=c(1, .87, .87, .75, .75, 1)),
    `16`=data.frame(x=c(.73, .73, 1, 1), y=c(.75, .475, .475, .75)),
    `17`=data.frame(x=c(.65, .65, 1, 1, .73, .73), y=c(.28, 0, 0, .475, .475, .28)),
    `18`=data.frame(x=c(.35, .35, .65, .65), y=c(.28, 0, 0, .28)),
    `19`=data.frame(x=c(0, 0, .35, .35, .27, .27), y=c(.475, 0, 0, .28, .28, .475)),
    `20`=data.frame(x=c(0, 0, .27, .27), y=c(.75, .475, .475, .75)),
    `21`=data.frame(x=c(0, 0, .27, .27, .35, .35), y=c(1, .75, .75, .87, .87, 1))
)

dat <- lapply(dat, function(x) {
    x$order <- 1:nrow(x)
    x
})

space.manikin.shape <- list_df2df(dat, "id")[, c(2, 3, 1, 4)]
```


And the code for the centers:

```{r}
centers <- data.frame(
    id = unique(space.manikin.shape$id),
    center.x=c(.5, .5, .625, .5, .375, .5, .66, .69, .66, .5, .34, .31, 
        .34, .5, .79, .815, .79, .5, .16, .135, .16),
    center.y=c(.597, .71, .5975, .5, .5975, .82, .81, .5975, .39, .3775, .39, 
        .5975, .81, .935, .89, .6025, .19, .14, .19, .6025, .89)
)

```

There you have it folks your very own quasi-shape file.  Celebrate the fruits of your labor by plotting that bad oscar.

---

## 2. Plot Your Homemade Quasi-Shape File

```{r, fig.width=4.5, fig.height=4.5}
ggplot(centers) + annotation_custom(img,0,1,0,1) +
    geom_map(aes(map_id = id), map = space.manikin.shape, colour="black", fill=NA) +
    theme_bw()+ 
    expand_limits(space.manikin.shape) +
    geom_text(data=centers, aes(center.x, center.y, label = id), color="grey60") 
```
`r caption_figure("fig4", "Plotting the Quasi-Shape File and a Raster Image")`

Then I said I may want to tone down the color of the silhouette a bit so I can plot geoms atop without distraction.  Here's that attempt.

```{r, fig.width=4.5, fig.height=4.5}
img[["raster"]][img[["raster"]] == "#0E0F0FFF"] <- "#E7E7E7"

ggplot(centers) + annotation_custom(img,0,1,0,1) +
    geom_map(aes(map_id = id), map = space.manikin.shape, colour="black", fill=NA) +
    theme_bw()+ 
    expand_limits(space.manikin.shape) +
    geom_text(data=centers, aes(center.x, center.y, label = id), color="grey60") 
```
`r caption_figure("fig5", "Altered Raster Image Color")`

---

##  3. Add an External svg/ps    

I realized quickly a raster was messy.  I read up a bit on them in the R Journal ([click here](http://journal.r-project.org/archive/2011-1/RJournal_2011-1_Murrell.pdf)).  In the process of reading and fooling around with [Picasa](http://picasa.google.com/) I turned my original silhouette (body.png) blue and couldn't fix him.  I headed back to http://www.flaticon.com/free-icon/standing-frontal-man-silhouette_10633 to download another.  In this act I saw you could download a svg file of the silhouette.  I thought maybe this will be less messier and easier to change colors.  This lead me to a google search and finding the `grImport` package after seeing this [listserve post](https://stat.ethz.ch/pipermail/r-sig-geo/2009-February/005003.html).  And then I saw an article from Paul `r citet(bib[["Murrell2009"]])` and figured I could turn the svg (I didn't realize what svg was until I opened it in Notepad++) into a ps file and read into R and convert to a flexible grid grob.

Probably there are numerous ways to convert an svg to a ps file but I chose a [cloud convert service](https://cloudconvert.org/svg-to-ps).  After I read the file in with `grImport` per the Paul `r citet(bib[["Murrell2009"]])` article.  You're going to have to download the ps file [HERE](https://github.com/trinker/space_manikin/raw/master/images/being.ps) and get to your working directory.


```{r, fig.width=4.5, fig.height=4.5, results='hide'}
browseURL("https://github.com/trinker/space_manikin/raw/master/images/being.ps")
## Move that file from your downloads to your working directory.
## Sorry I don't know how to automate this.
library(grImport)

## Convert to xml
PostScriptTrace("being.ps")

## Read back in and convert to a grob
being_img <- pictureGrob(readPicture("being.ps.xml"))

## Plot it
ggplot(centers) + annotation_custom(being_img,0,1,0,1) +
    geom_map(aes(map_id = id), map = space.manikin.shape, 
    	colour="black", fill=NA) +
    theme_bw()+ 
    expand_limits(space.manikin.shape) +
    geom_text(data=centers, aes(center.x, center.y, 
    	label = id), color="grey60") 
```
`r caption_figure("fig6", "Quasi-Shape File with Grob Image Rather than Raster")`

---

## 4. Change a `grid` Grob's Color and Alpha   

Now we have a flexible grob we can mess around with colors and alpha until our heart's content.

`str` is our friend to figure out where and how to mess with the grob (`str(being_img)`).  That leads me to the following changes to the image to adjust color and/or alpha (transparency).

```{r, fig.width=4.5, fig.height=4.5}
being_img[["children"]][[1]][[c("gp", "fill")]] <- 
  being_img[["children"]][[2]][[c("gp", "fill")]] <- "black"

being_img[["children"]][[1]][[c("gp", "alpha")]] <- 
  being_img[["children"]][[2]][[c("gp", "alpha")]] <- .2

## Plot it
ggplot(centers) + annotation_custom(being_img,0,1,0,1) +
    geom_map(aes(map_id = id), map = space.manikin.shape, 
    	colour="black", fill=NA) +
    theme_bw()+ 
    expand_limits(space.manikin.shape) +
    geom_text(data=centers, aes(center.x, center.y, 
    	label = id), color="grey60") 
```
`r caption_figure("fig6", "Quasi-Shape File with Grob Image Alpha = .2")`

---

## Let's Have Some Fun

Let's make it into a choropleth and a density plot.  We'll make some fake fill values to fill with.

```{r, fig.width=6.5, fig.height=5}
set.seed(10)
centers[, "Frequency"] <- rnorm(nrow(centers))
	
being_img[["children"]][[1]][[c("gp", "alpha")]] <- 
  being_img[["children"]][[2]][[c("gp", "alpha")]] <- .25

ggplot(centers, aes(fill=Frequency)) +
    geom_map(aes(map_id = id), map = space.manikin.shape, 
    	colour="black") +
	scale_fill_gradient2(high="red", low="blue") +
    theme_bw()+ 
    expand_limits(space.manikin.shape) +
    geom_text(data=centers, aes(center.x, center.y, 
    	label = id), color="black") + 
	annotation_custom(being_img,0,1,0,1) 
```
`r caption_figure("fig7", "Quasi-Shape File as a Choropleth")`

```{r, fig.width=6.5, fig.height=5}
set.seed(10)
centers[, "Frequency2"] <- sample(seq(10, 150, by=20, ), nrow(centers), TRUE)

centers2 <- centers[rep(1:nrow(centers), centers[, "Frequency2"]), ]

ggplot(centers2) +
#	    geom_map(aes(map_id = id), map = space.manikin.shape, 
#    	colour="grey65", fill="white") +
    stat_density2d(data = centers2, 
        aes(x=center.x, y=center.y, alpha=..level.., 
        fill=..level..), size=2, bins=12, geom="polygon") + 
    scale_fill_gradient(low = "yellow", high = "red") +
    scale_alpha(range = c(0.00, 0.5), guide = FALSE) +
    theme_bw()+ 
    expand_limits(space.manikin.shape) +
    geom_text(data=centers, aes(center.x, center.y, 
    	label = id), color="black") + 
	annotation_custom(being_img,0,1,0,1) +
    geom_density2d(data = centers2, aes(x=center.x, 
        y=center.y), colour="black", bins=8, show_guide=FALSE) 
```
`r caption_figure("fig8", "Quasi-Shape File as a Density Plot")`
	
	
Good times were had by all.

---

<em><font size="2">Created using the `r HR2("https://github.com/trinker/reports", "reports")` `r citep(bib["Rinker2013b"])` package</font></em>

<font size="2">Get the .Rmd file `r HR2("https://github.com/trinker/space_manikin/blob/master/space_manikin.Rmd", "here")`</font>

--- 
## References
```{r, echo=FALSE, results='asis'}
bibliography("html")
```

```{r, echo=FALSE, results="hide"}
knitr::purl("space_manikin.Rmd")
```
